---
title: "Education level and employment status data analysis"
author: "Emma Wagner"
date: "11/15/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this R markdown, I will analyze the education and employment dataset. 
To start, I loaded the necessary libraries and read in the excel sheet. I used the head() function as well to make sure the data was read in correctly.
```{r}
library(tidyverse)
library(readxl)

edu = read_excel("/Users/emmawagner/Desktop/Data Wrangling/Final project/Mental health data cleaned.xlsx", sheet = "depression-edu-cleaned")
head(edu)
```

Next, I created a preliminary bar plot to look at the data. I created a bar graph with country on the x axis and the depression rate (%) on the y axis. 
```{r}
bar = ggplot(edu, aes(x= Code, y = All_total)) +
  geom_bar(stat= "identity") +
  theme(text = element_text(size=8))
bar
```

Here we can see there is a range in depression rates between different countries- however this is not very useful for analyses of education and employment. In order to better visualize and analyze the data using the tidyverse, we need to convert the data into long form.  

```{r}
# Put data into long format
edu_long = pivot_longer(edu, cols = c("All_active", "All_employed", "All_total", "BelowUS_active", "BelowUS_employed", "BelowUS_total", 
                        "Tertiary_active", "Tertiary_employed", "Tertiary_total", "US_active", "US_employed", "US_total"),
                        names_to = "EduLevel")

# Split EduLevel into two columns- EduLevel and Employment
edu_long = separate(edu_long, col = EduLevel, into=c("EduLevel", "Employment"), sep = "_")

```
edu_long now contains the data in long form, with three new columns. EduLevel and Employment contains the education level and employment connected to each depression rate. The value column contains the rate of depression. This reduces the number of variables to 6, from 15 previously.

Next, I created a barplot of data from all of the countries to do some initial analysis of education. I filtered the data to include only the total data for employment. 
```{r}
# barplot by countries
level_order = c("All", "BelowUS", "US", "Tertiary")
totalonly= edu_long %>%
  filter(Employment == "total")
bar = ggplot(totalonly, aes(x= Code, y = value, fill = factor(EduLevel, level= level_order))) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  scale_fill_brewer(palette = "Paired", labels = c("All", "Below upper secondary", "Upper secondary", "Tertiary")) +
  theme(text = element_text(size=6)) + 
  labs(y= "Rate of depresssion (%)", x = "Country", fill = "Education level") 
bar
```
In every country, the highest rates of depression were among people with a below upper secondary level of education and the lowest rates of depression were for people with a tertiary level education. This indicates that education reduces depression rates. 

Next, I did the same thing for employment data, filtering to include all levels of education. 
```{r}
# barplot by countries
level_order = c("total", "active", "employed")
allonly= edu_long %>%
  filter(EduLevel == "All")
bar = ggplot(allonly, aes(x= Code, y = value, fill = factor(Employment, level= level_order))) +
  geom_bar(stat = "identity", position = "dodge") +
  theme_classic() +
  scale_fill_brewer(palette = "Paired", labels = c("Total", "Actively looking", "Employed")) +
  theme(text = element_text(size=6)) + 
  labs(y= "Rate of depresssion (%)", x = "Country", fill = "Employment status") 
bar
```
Every country follows the same pattern- employed people have the lowest rates of depression, people actively looking for work have a slightly higher rate, and the total population (including unemployed people) had the highest rates. This supports the idea that employment reduces rates of depression. 

In these two graphs, there is a significant level of variation in depression rates between countries. My group mates analyzed depression prevalence at the regional and country level with bigger data sets, so I am not going to focus on differences between countries, just on differences between groups based on education and employment. 

With 26 countries the above visualizations are a bit cluttered, so in the next plots I will aggregate the data.

Below, I filtered the data by employment level to create plots of the average depression rate for that group. Each plot shows the depression rates for different levels of education for that employment status.
```{r}
level_order = c("All", "BelowUS", "US", "Tertiary")
# for total data
bar2 =  edu_long %>%
  filter(Employment == "total") %>%
  group_by(EduLevel) %>%
  summarise(meanByEdu = mean(value)) %>%
  ggplot(aes(x= factor(EduLevel, level= level_order), y = meanByEdu, fill = EduLevel)) +
  geom_col() +
  scale_y_continuous(limits =c(0, 12.5)) +
  theme_classic() +
  geom_text(aes(label = (round(meanByEdu, digits = 2)), vjust = -0.5)) +
  labs(x= "Education level", y= "Average rate of depression (%)", title = "All employment statuses") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_discrete(labels = c("All", "Below upper \n secondary", "Upper secondary", "Tertiary")) 
bar2

# for active data
bar3 =  edu_long %>%
  filter(Employment == "active") %>%
  group_by(EduLevel) %>%
  summarise(meanByEdu = mean(value)) %>%
  ggplot(aes(x= factor(EduLevel, level= level_order), y = meanByEdu, fill = EduLevel)) +
  geom_col() +
  scale_y_continuous(limits =c(0, 12.5)) +
  geom_text(aes(label = (round(meanByEdu, digits = 2)), vjust = -0.5)) +
  theme_classic() +
  labs(x= "Education level", y= "Average rate of depression (%)", title = "Actively looking for employment") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_discrete(labels = c("All", "Below upper \n secondary", "Upper secondary", "Tertiary")) 
bar3

# for employed
bar3 =  edu_long %>%
  filter(Employment == "employed") %>%
  group_by(EduLevel) %>%
  summarise(meanByEdu = mean(value)) %>%
  ggplot(aes(x= factor(EduLevel, level= level_order), y = meanByEdu, fill = EduLevel)) +
  geom_col() +
  scale_y_continuous(limits =c(0, 12.5)) +
  geom_text(aes(label = (round(meanByEdu, digits = 2)), vjust = -0.5)) +
  theme_classic() +
  labs(x= "Education level", y= "Average rate of depression (%)", title = "Employed") +
  theme(legend.position = "none") +
  scale_fill_brewer(palette = "Paired") +
  scale_x_discrete(labels = c("All", "Below upper \n secondary", "Upper secondary", "Tertiary")) 
bar3
```
These plots show the same patterns as the country level plots above: across all levels of employment, the more educated groups had lower average rates of depression than less educated groups. Similarly, employed people had lower average rates of depression than both those actively looking for work and the total population. Overall, the lowest average rate of depression (4.33%) was for the employed and tertiary education groups. The highest average rate (12.28%) was for below upper secondary education in the total population.

For a final set of plots, I plotted scatterplots with smoothing lines to show the trends across countries for another visual display. Each plot is for a different employment status
```{r}
# Scatter plot with smoothing lines for total employment data
all =edu_long %>%
  filter(Employment == "total")
p = ggplot(all, aes(x = Code, y = value, fill = EduLevel, color= EduLevel, group = EduLevel)) +
  geom_point()+
  geom_smooth(method = loess, se =FALSE) +
  scale_color_brewer(palette = "Paired") +
  theme_classic() +
  theme(text = element_text(size=6)) + 
  labs(title = "Total")
p

# Scatter plot with smoothing lines for active
active =edu_long %>%
  filter(Employment == "active")
p = ggplot(active, aes(x = Code, y = value, fill = EduLevel, color= EduLevel, group = EduLevel)) +
  geom_point()+
  geom_smooth(method = loess, se =FALSE) +
  scale_color_brewer(palette = "Paired") +
  theme_classic() + 
  theme(text = element_text(size=6)) + 
  labs(title = "Actively looking for employment")
p

# Scatter plot with smoothing lines for employed
employ =edu_long %>%
  filter(Employment == "employed")
p = ggplot(employ, aes(x = Code, y = value, fill = EduLevel, color= EduLevel, group = EduLevel)) +
  geom_point()+
  geom_smooth(method = loess, se =FALSE) +
  scale_color_brewer(palette = "Paired") +
  theme_classic() +
  theme(text = element_text(size=6)) + 
  labs(title = "Employed")
p

```
Overall, this showed the same trends as the plots above. More educated people had lower levels of depression and less educated people had higher rates. Because of the amount of variation between countries, this plot is not quite as effective at visualizing the data as the bar plots above.

Next, I ran some statistical tests on the data to see if the differences between groups are statistically significant.

First, I ran a one-way anova to see if there were significant differences between groups based on education levels. 
```{r}
summary(aov(value~EduLevel, data = edu_long))

```
The p value< 0.05, so the test is statistically significant and we can reject the null hypothesis of no differences between groups. 

Next, I ran a one way anova based on employment
```{r}
summary(aov(value~Employment, data = edu_long))

```
Again, the p value is less than .05 we can reject the null hypothesis, indicating that the differences between groups are statistically significant.

Finally, I ran a two way anova with both EduLevel and Employment.
```{r}
summary(aov(value~Employment +EduLevel, data = edu_long))
```
Both p values are less than 0.05 indicating the differences are statistically significant.

Overall, the data showed clearly the positive effects of education and employment on reducing depression rates. This highlights how issues such as the increasing unaffordability of college, unemployment, and lack of quality public school educations for all students are detrimental to mental health at the population level.

In future analyses, it would be interesting to analyze a dataset with more countries to see if these trends hold true in other countries and regions. Initially, I was not expecting education to reduce depression rates so significantly. In the United States, there is a pretty unhealthy culture around work with long hours with little vacation time. So, I was expecting that people with higher levels of education would be more likely to have high stress jobs and would have rates of depression similar to those with less education. On the other hand, in Europe, where this data is from, there seems to be a more balanced culture around work, with more time off, paid parental leave, etc. I'm curious if this impacts depression rates. So, it would be interesting to analyze how culture and policies around employment and education influence depression rates. Additionally, this data was only from 2014. Analyzing data from other years could allow for more complex analyses and allow us to analyze trends over time.